A comparison of design-based and model-based approaches for spatial data.

### 2021-08-04 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
wrote functions to allow a more easy comparison between grid and random locations, and added simulation thoughts to manuscript.

>>>>

```{r, results = "hide", message = FALSE, warning = FALSE}
library(tidyverse)
files <- list.files(path = here("inst", "output"), pattern = "*.csv",
                    full.names = TRUE)

combo_data <- purrr::map_df(files,
                            ~read_csv(.x) %>% mutate(filename = .x)) %>%
  separate(filename, into = c("junk", "error_dep", "dist", "n", "location"),
           sep = "_") %>%
  select(-junk)
combo_data <- combo_data %>% mutate(sim = interaction(error_dep, dist, n, location))

combo_data_grts <- combo_data %>% filter(approach %in% c("Design GRTS", "Model GRTS"))

dep_only <- combo_data_grts %>% filter(error_dep == "deperr")
dep_long <- dep_only %>%
  pivot_wider(names_from = approach, values_from = rmspe, sim) %>%
  mutate(perc = 100 * (`Design GRTS` - `Model GRTS`) / `Model GRTS`)
```

Major Points from August 3 Simulations:

1. In most of the dependent error simulation settings, either all four approahces (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model) perform equally or GRTS-Design and GRTS-Model outperform IRS-Design and IRS-Model. Exceptions to this are a couple of the settings with very small sample sizes (n = 10), in which the IRS does better than GRTS. In the independent error settings, it usually doesn't matter much which approach is used, which makes sense.

```{r}
ggplot(data = combo_data, aes(x = sim, y = rmspe, colour = approach)) +
  geom_point() +
  coord_flip()
```

2. We will now focus in a bit more on comparing Design-GRTS to Model-GRTS, the two best approaches for any reasonable sample size. In the independent error settings, the two approaches perform very similarly, so those results are omitted in the following graph. In the dependent error settings, using rmspe as the performance criterion, Model-GRTS outperforms Design-GRTS in 12 of the 18 settings, the two approaches perform very similarly in 3 settings, and Design-GRTS outperforms Model-GRTS in 3 settings.

```{r}
ggplot(data = dep_long, aes(x = fct_reorder(sim, perc), y = perc)) +
  geom_point() +
  coord_flip() +
  labs(y = "Percent Change in rmspe (Design - Model) / Model",
       caption = "Percent > 0 means Model has better rmspe")
```

3. Focusing in on the three settings where Design-GRTS outperforms Model-GRTS, we see that, in two of the settings, the log-normal response has a large variance, corresponding to a large right-skew after exponentiation. All three settings have sites in random locations. However, in only one of these settings would we recommend actually using Design-GRTS. In the other two settings, the data are sufficiently skewed that a practitioner should not use either approach, though it is "safer" to use Design-GRTS.

```{r}
design_better <- dep_long %>% filter(perc < -5)
design_better_full <- inner_join(combo_data_grts, design_better, by = c("sim"))
ggplot(data = design_better_full, aes(x = sim, y = coverage, colour = approach)) +
  geom_point() +
  coord_flip() +
  geom_hline(aes(yintercept = 0.95), alpha = 0.5)
```

4. Take-home messages

- In terms of rmspe, a model-based analysis on a GRTS design yields an rmspe similar to or lower than a design-based analysis on a GRTS design, as long as the response variable is not "too skewed." 
- If the response variable is very skewed, then neither analysis is appropriate, but, the design-based analysis is quicker. 
- a spatially balanced GRTS sample outperforms IRS in nearly all dependent error settings, as expected.
- methods that use spatial correlation generally perform better on random location points than they do on gridded points. This makes some intuitive sense because (1) on average, the minimum distance between an unobserved point and its nearest observed neighbor should be lower for random points and (2) the span of the study area is maximized for a grid based on the way that we set up the simulations (with the random points being drawn as uniform random variables within the boundary of the grid). 
- comparison of Design-GRTS and Model-GRTS between two settings with different locations of points, but otherwise the same simulation parameters, should really be done on the same surface realization. One very strange realized response vector could drastically alter the results, especially on the exponentiated log data. In the same way that we compare the four approaches on the same realized data, we should also try to do the same with the locations, if they are of interest. (The realized mean won't be exactly the same but should be close).

### 2021-08-04 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
wrote functions to allow a more easy comparison between grid and random locations, and added simulation thoughts to manuscript.

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A comparison of design-based and model-based approaches for spatial data.},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A comparison of design-based and model-based approaches for spatial data.},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

Major Points from August 3 Simulations:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  In most of the dependent error simulation settings, either all four
  approahces (IRS-Design, IRS-Model, GRTS-Design, and GRTS-Model)
  perform equally or GRTS-Design and GRTS-Model outperform IRS-Design
  and IRS-Model. Exceptions to this are a couple of the settings with
  very small sample sizes (n = 10), in which the IRS does better than
  GRTS. In the independent error settings, it usually doesn't matter
  much which approach is used, which makes sense.
\end{enumerate}

\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  We will now focus in a bit more on comparing Design-GRTS to
  Model-GRTS, the two best approaches for any reasonable sample size. In
  the independent error settings, the two approaches perform very
  similarly, so those results are omitted in the following graph. In the
  dependent error settings, using rmspe as the performance criterion,
  Model-GRTS outperforms Design-GRTS in 12 of the 18 settings, the two
  approaches perform very similarly in 3 settings, and Design-GRTS
  outperforms Model-GRTS in 3 settings.
\end{enumerate}

\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Focusing in on the three settings where Design-GRTS outperforms
  Model-GRTS, we see that, in two of the settings, the log-normal
  response has a large variance, corresponding to a large right-skew
  after exponentiation. All three settings have sites in random
  locations. However, in only one of these settings would we recommend
  actually using Design-GRTS. In the other two settings, the data are
  sufficiently skewed that a practitioner should not use either
  approach, though it is ``safer'' to use Design-GRTS.
\end{enumerate}

\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Take-home messages
\end{enumerate}

\begin{itemize}
\tightlist
\item
  In terms of rmspe, a model-based analysis on a GRTS design yields an
  rmspe similar to or lower than a design-based analysis on a GRTS
  design, as long as the response variable is not ``too skewed.''
\item
  If the response variable is very skewed, then neither analysis is
  appropriate, but, the design-based analysis is quicker.
\item
  a spatially balanced GRTS sample outperforms IRS in nearly all
  dependent error settings, as expected.
\item
  methods that use spatial correlation generally perform better on
  random location points than they do on gridded points. This makes some
  intuitive sense because (1) on average, the minimum distance between
  an unobserved point and its nearest observed neighbor should be lower
  for random points and (2) the span of the study area is maximized for
  a grid based on the way that we set up the simulations (with the
  random points being drawn as uniform random variables within the
  boundary of the grid).
\item
  comparison of Design-GRTS and Model-GRTS between two settings with
  different locations of points, but otherwise the same simulation
  parameters, should really be done on the same surface realization. One
  very strange realized response vector could drastically alter the
  results, especially on the exponentiated log data. In the same way
  that we compare the four approaches on the same realized data, we
  should also try to do the same with the locations, if they are of
  interest. (The realized mean won't be exactly the same but should be
  close).
\end{itemize}
